---
title: "CreditCardDefault"
author: "Rajkumar Conjeevaram Mohan"
date: "28/04/2022"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r load_data}
data.orig <- read_xls(path='default of credit card clients.xls',
                      sheet='Data',  skip=1)
data.orig <- data.frame(data.orig)
colnames(data.orig) <- c('id', 'total_credit', 'gender', 'education', 'marital', 'age', 'rs_sep', 'rs_aug', 'rs_july', 'rs_june', 'rs_may', 'rs_apr', 'bs_sep', 'bs_aug', 'bs_july', 'bs_june', 'bs_may', 'bs_apr', 'ap_sep', 'ap_aug', 'ap_july', 'ap_june', 'ap_may', 'ap_apr', 'default_r')

data.orig$marital <- factor(data.orig$marital)
data.orig$gender <- factor(data.orig$gender)
data.orig$education <- factor(data.orig$education)
data.orig$rs_sep <- factor(data.orig$rs_sep)
data.orig$rs_aug <- factor(data.orig$rs_aug)
data.orig$rs_july <- factor(data.orig$rs_july)
data.orig$rs_june <- factor(data.orig$rs_june)
data.orig$rs_may <- factor(data.orig$rs_may)
data.orig$default_r <- factor(data.orig$default_r)
# data.orig
```


# Data Description
```{r}
str(data.orig)
```

# Data overview
```{r, results='markup'}

```

# Data Imbalance
```{r}
ggplot(data.orig, aes(x=default_r)) + 
geom_bar() +
xlab('Credit Card Default') +
ggtitle('Imbalanced distribution of response variable')  
```

# EDA

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



